---
title: "1_PLSC_grad_behav"
author: "Ju-Chi.Yu"
date: "3/25/2022"
output: html_document
---

<style type="text/css">
.main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
</style>

Grabbing SPINS gradients

```{r, echo = FALSE}
library(tidyverse)
library(ggseg)
# library(ggsegGlasser)
# library(broom)
library(TExPosition)
library(PTCA4CATA)
library(plotly)
library(colorspace)
library("RColorBrewer")
```

## Read in the SPINS big table


```{r, echo = FALSE, warning= FALSE, message='hide'}

spins_grads_wGSR <- read_csv("../gsr_spins_concat_full")
# spins_grads <- spins_grads %>%
#   mutate(site = str_sub(`Subject`, 5, 7)) %>%
#   rename(task = `EA or RS`, 
#          subject = `Subject ID`)
spins_grads_noGSR <- read_csv("../spins_concat_full")

spins_grads <- bind_rows("GSR" = spins_grads_wGSR,
                         "noGSR" = spins_grads_noGSR,
                         .id = "GSR")

rm(spins_grads_noGSR, spins_grads_wGSR)

spins_grads_num <- data.frame(spins_grads[,c(3:5, 13,15)])
spins_grads_num_full <- data.frame(spins_grads[,c(3:5, 13,15,16)])
spins_grads_wide <- reshape(spins_grads_num, idvar = "Subject", timevar = "ROI", direction = "wide")
```

## read subject data

```{r, echo = FALSE, warning= FALSE, message='hide'}
lol_spins_behav <- 
  read_csv('../data/spins_lolivers_subject_info_for_grads_2022-02-02.csv') %>%
  filter(exclude_MRI==FALSE, 
         has_EA_grads==TRUE, 
         has_RS_grads==TRUE)
names(lol_spins_behav)

## design matrix for subjects
spins_dx <- lol_spins_behav %>%
  select(subject,site,scanner,diagnostic_group,demo_sex,demo_age_study_entry)
spins_dx_org <- spins_dx[,-1] %>% data.frame
rownames(spins_dx_org) <- spins_dx$subject

## numeric data
spins_behav_num <- lol_spins_behav %>% 
  select(scog_tasit1_total, scog_tasit2_total, scog_tasit3_total, np_composite_tscore, 
np_domain_tscore_process_speed,	np_domain_tscore_att_vigilance,	np_domain_tscore_work_mem,	np_domain_tscore_verbal_learning,	np_domain_tscore_visual_learning,	np_domain_tscore_reasoning_ps,	np_domain_tscore_social_cog,	bprs_factor_pos_symp,	bprs_factor_total,	bprs_factor_neg_symp,	qls_total,	sans_total_sc,	bsfs_total
  ) %>% data.frame
colnames(spins_behav_num)
rownames(spins_behav_num) <- lol_spins_behav$subject

## participants' variables of which the effects should be regressed out

var2regout <- lol_spins_behav %>%
  select(demo_sex, demo_age_study_entry, `fd_mean_task-rest`) %>% data.frame
rownames(var2regout) <- lol_spins_behav$subject
var2regout$demo_sex_num <- as.numeric(as.factor(var2regout$demo_sex))-1
var2regout_num <- var2regout[,-1]
```

## Check subject overlap

```{r, warning= FALSE, message='hide'}
grad.sub <- spins_grads_wide$Subject[order(spins_grads_wide$Subject)]
behav.sub <- lol_spins_behav$subject[order(lol_spins_behav$subject)]

behav.sub[behav.sub %in% grad.sub == FALSE]
grad.sub[grad.sub %in% behav.sub == FALSE]

## kept subjects
kept.sub <- behav.sub[behav.sub %in% grad.sub] # 342
kept.sub[complete.cases(spins_behav_num[kept.sub,c(1:11, 17)]) == FALSE]
kept.sub <- kept.sub[complete.cases(spins_behav_num[kept.sub,c(1:11, 17)])] # 338

## grab the matching data

behav.dat <- spins_behav_num[kept.sub,c(1:11, 17)]
spins_grads_wide_org <- spins_grads_wide[,-1]
rownames(spins_grads_wide_org) <- spins_grads_wide$Subject
grad.dat <- spins_grads_wide_org[kept.sub,]

## variables to regress out
regout.dat <- var2regout_num[kept.sub,]

rm(spins_grads_wide, spins_grads_wide_org, lol_spins_behav, spins_behav_num, spins_grads_num)

```

## Regress out the effects

```{r}
behav.reg <- apply(behav.dat, 2, function(x) lm(x~regout.dat$demo_sex + regout.dat$demo_age_study_entry + regout.dat$fd_mean_task.rest)$residual)

grad.reg <- apply(grad.dat, 2, function(x) lm(x~regout.dat$demo_sex + regout.dat$demo_age_study_entry + regout.dat$fd_mean_task.rest)$residual)
```

## grab some network colours

```{r, warning= FALSE, message='hide'}
networks <- read_delim("../networks.txt", 
                       "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  select(NETWORK, NETWORKKEY, RED, GREEN, BLUE, ALPHA) %>%
  distinct() %>%
  add_row(NETWORK = "Subcortical", NETWORKKEY = 13, RED = 0, GREEN=0, BLUE=0, ALPHA=255) %>%
  mutate(hex = rgb(RED, GREEN, BLUE, maxColorValue = 255)) %>%
  arrange(NETWORKKEY)

networks$hex <- darken(networks$hex, 0.2)

# oi <- networks$hex
# swatchplot(
#   "-40%" = lighten(oi, 0.4),
#   "-20%" = lighten(oi, 0.2),
#   "  0%" = oi,
#   " 20%" =  darken(oi, 0.2),
#   " 25%" =  darken(oi, 0.25),
#   " 30%" =  darken(oi, 0.3),
#   " 35%" =  darken(oi, 0.35),
#   off = c(0, 0)
# )

networks
```

## get row and column designs

```{r}
## match ROIs to networks
ROI.network.match <- cbind(spins_grads$ROI, spins_grads$Network) %>% unique
ROI.idx <- ROI.network.match[,2]
names(ROI.idx) <- ROI.network.match[,1]
### match networks with colors
net.col.idx <- networks$hex
names(net.col.idx) <- networks$NETWORK

## design matrix for subjects
sub.dx <- spins_dx_org[kept.sub,]

diagnostic.col <- sub.dx$diagnostic_group %>% as.matrix %>% makeNominalData() %>% createColorVectorsByDesign()
rownames(diagnostic.col$gc) <- sub(".","", rownames(diagnostic.col$gc))

## design matrix for columns - behavioral 
behav.dx <- matrix(nrow = ncol(behav.dat), ncol = 1, dimnames = list(colnames(behav.dat), "type")) %>% as.data.frame

behav.col <- c("scog" = "#F28E2B",
               "np" = "#59A14F",
               "bsfs" = "#D37295")

behav.dx$type <- sub("(^[^_]+).*", "\\1", colnames(behav.dat))
behav.dx$type.col <- recode(behav.dx$type, !!!behav.col)

## design matrix for columns - gradient
grad.dx <- matrix(nrow = ncol(grad.dat), ncol = 4, dimnames = list(colnames(grad.dat), c("gradient", "ROI", "network", "network.col"))) %>% as.data.frame

grad.dx$gradient <- sub("(^[^.]+).*", "\\1", colnames(grad.dat))
grad.dx$ROI <- sub("^[^.]+.(*)", "\\1", colnames(grad.dat))
grad.dx$network <- recode(grad.dx$ROI, !!!ROI.idx)
grad.dx$network.col <- recode(grad.dx$network, !!!net.col.idx)

## get different alpha for gradients
grad.col.idx <- c("grad1" = "grey30",
                  "grad2" = "grey60",
                  "grad3" = "grey90")
grad.dx$gradient.col <- recode(grad.dx$gradient, !!!grad.col.idx)

## for heatmap
col.heat <- colorRampPalette(c("red", "white", "blue"))(256)
```

## Run PLS-C

```{r}
pls.res <- tepPLS(behav.reg, grad.reg, DESIGN = sub.dx$diagnostic_group, make_design_nominal = TRUE, graphs = FALSE)

pls.boot <- data4PCCAR::Boot4PLSC(behav.reg, grad.reg, scale1 = TRUE, scale2 = TRUE)
## swith direction for dimension 3
# pls.res$TExPosition.Data$fi[,3] <- pls.res$TExPosition.Data$fi[,3]*-1
# pls.res$TExPosition.Data$fj[,3] <- pls.res$TExPosition.Data$fj[,3]*-1
# pls.res$TExPosition.Data$pdq$p[,3] <- pls.res$TExPosition.Data$pdq$p[,3]*-1
# pls.res$TExPosition.Data$pdq$q[,3] <- pls.res$TExPosition.Data$pdq$q[,3]*-1
# pls.res$TExPosition.Data$lx[,3] <- pls.res$TExPosition.Data$lx[,3]*-1
# pls.res$TExPosition.Data$ly[,3] <- pls.res$TExPosition.Data$ly[,3]*-1

## Scree plot
PlotScree(pls.res$TExPosition.Data$eigs)

## Print singular values
pls.res$TExPosition.Data$pdq$Dv
## Print eigenvalues
pls.res$TExPosition.Data$eigs
## Compare the inertia to the largest possible inertia
sum(cor(behav.dat, grad.dat)^2)
sum(cor(behav.dat, grad.dat)^2)/(ncol(behav.dat)*ncol(grad.dat))
```
Here, we show that the effect that PLSC decomposes is pretty small to begin with. The effect size of the correlation between the two tables is 92.40 which accounts for 0.0065 of the largest possible effect. 

## Results {.tabset}

```{r, include = FALSE}
lxly.out <- list()

for (check.dim in 1:3){
  
  lxly <- cbind(pls.res$TExPosition.Data$lx[,check.dim], pls.res$TExPosition.Data$ly[,check.dim])
  colnames(lxly) <- c(paste0("Dim", check.dim, c(".Behavioural", ".gradient")))
  
  lxly.boot <- Boot4Mean(lxly, sub.dx$diagnostic_group, niter = 1000)
  colnames(lxly.boot$GroupMeans) <- colnames(lxly.boot$BootCube) <- c(paste0("Dim", check.dim, c(".Behavioural", ".gradient")))
  
  ## plot latent variables
  lxly.all <- createFactorMap(lxly,
                              title = paste0("Latent variables"),
                              col.background = NULL,
                              col.axes = "orchid4",
                              alpha.axes = 0.5,
                              col.points = diagnostic.col$oc,
                              alpha.points = 0.1)
  
  lxly.avg <- createFactorMap(lxly.boot$GroupMeans,
                              col.points = diagnostic.col$gc[rownames(lxly.boot$GroupMeans),],
                              col.labels =  diagnostic.col$gc[rownames(lxly.boot$GroupMeans),], 
                              pch = 17, alpha.points = 1, text.cex = 5)
  
  lxly.CI <- MakeCIEllipses(lxly.boot$BootCube,
                            col =  diagnostic.col$gc[rownames(lxly.boot$BootCube),],
                            names.of.factors = c(paste0("Dim", check.dim, c(".Behavioural", ".gradient"))), alpha.ellipse = 0.1, line.size = 0.5)
  
  lxly.out[[check.dim]] <- lxly.all$zeMap_background + lxly.all$zeMap_dots + lxly.CI + lxly.avg$zeMap_dots + lxly.avg$zeMap_text + coord_cartesian()
  
}

q.grad1 <- pls.res$TExPosition.Data$fj[which(grad.dx$gradient == "grad1"),]
q.sig.grad1 <- pls.boot$bootRatiosSignificant.j[which(grad.dx$gradient == "grad1"),]
col.grad1 <- grad.dx$network.col[which(grad.dx$gradient == "grad1")]
dx.grad1 <- grad.dx$network[which(grad.dx$gradient == "grad1")]

q.grad2 <- pls.res$TExPosition.Data$fj[which(grad.dx$gradient == "grad2"),]
q.sig.grad2 <- pls.boot$bootRatiosSignificant.j[which(grad.dx$gradient == "grad2"),]
col.grad2 <- grad.dx$network.col[which(grad.dx$gradient == "grad2")]
dx.grad2 <- grad.dx$network[which(grad.dx$gradient == "grad2")]

q.grad3 <- pls.res$TExPosition.Data$fj[which(grad.dx$gradient == "grad3"),]
q.sig.grad3 <- pls.boot$bootRatiosSignificant.j[which(grad.dx$gradient == "grad3"),]
col.grad3 <- grad.dx$network.col[which(grad.dx$gradient == "grad3")]
dx.grad3 <- grad.dx$network[which(grad.dx$gradient == "grad3")]

```

### Dimension 1

```{r,include = FALSE}

col4grad1 <- col.grad1[order(dx.grad1)]
bar.grad1 <- PrettyBarPlot2(q.grad1[order(dx.grad1),1],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad1[,1] == TRUE, col4grad1, "grey90"), 
                            font.size = 0, main = "Scores - gradient 1")

col4grad2 <- col.grad2[order(dx.grad2)]
bar.grad2 <- PrettyBarPlot2(q.grad2[order(dx.grad2),1],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad2[,1] == TRUE, col4grad2, "grey90"), 
                            font.size = 0, main = "Scores - gradient 2")

col4grad3 <- col.grad3[order(dx.grad3)]
bar.grad3 <- PrettyBarPlot2(q.grad3[order(dx.grad3),1],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad3[,1] == TRUE, col4grad3, "grey90"), 
                            font.size = 0, main = "Scores - gradient 3")

```


```{r, fig.show = 'hold', out.width = '50%'}
lxly.out[[1]]

gridExtra::grid.arrange(bar.grad1, bar.grad2, bar.grad3, ncol = 1)

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,1],
               threshold = 0, 
               color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,1] == TRUE, behav.dx$type.col, "grey90"),
               horizontal = FALSE, main = "Scores - behavioural")

cor.heat <- pls.res$TExPosition.Data$X %>% heatmap(col = col.heat)

## control
grad.dat.ctrl <- grad.dat[sub.dx$diagnostic_group == "control",]
behav.dat.ctrl <- behav.dat[sub.dx$diagnostic_group == "control",]
corX.ctrl <- cor(as.matrix(behav.dat.ctrl),as.matrix(grad.dat.ctrl))
heatmap(corX.ctrl[cor.heat$rowInd, cor.heat$colInd], col = col.heat, Rowv = NA, Colv = NA)

## case
grad.dat.case <- grad.dat[sub.dx$diagnostic_group == "case",]
behav.dat.case <- behav.dat[sub.dx$diagnostic_group == "case",]
corX.case <- cor(as.matrix(behav.dat.case),as.matrix(grad.dat.case))
heatmap(corX.case[cor.heat$rowInd, cor.heat$colInd], col = col.heat, Rowv = NA, Colv = NA)

```

### Dimension 2

```{r, include = FALSE}
bar.grad1 <- PrettyBarPlot2(q.grad1[order(dx.grad1),2],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad1[,2] == TRUE, col4grad1, "grey90"), 
                            font.size = 0, main = "Scores - gradient 1")

bar.grad2 <- PrettyBarPlot2(q.grad2[order(dx.grad2),2],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad2[,2] == TRUE, col4grad2, "grey90"), 
                            font.size = 0, main = "Scores - gradient 2")

bar.grad3 <- PrettyBarPlot2(q.grad3[order(dx.grad3),2],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad3[,2] == TRUE, col4grad3, "grey90"), 
                            font.size = 0, main = "Scores - gradient 3")

```


```{r, fig.show = 'hold', out.width = '50%'}
lxly.out[[2]]

gridExtra::grid.arrange(bar.grad1, bar.grad2, bar.grad3, ncol = 1)

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,2],
               threshold = 0, color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,2] == TRUE, behav.dx$type.col, "grey90"), 
               horizontal = FALSE, main = "Scores - behavioural")

dim1.est <- pls.res$TExPosition.Data$pdq$Dv[1]*as.matrix(pls.res$TExPosition.Data$pdq$p[,1], ncol = 1) %*% t(as.matrix(pls.res$TExPosition.Data$pdq$q[,1], ncol = 1))


cor.heat.res1 <- (pls.res$TExPosition.Data$X - dim1.est) %>% heatmap(col = col.heat)

## control
res1.ctrl <- corX.ctrl - dim1.est
heatmap(res1.ctrl[cor.heat.res1$rowInd, cor.heat.res1$colInd], col = col.heat, Rowv = NA, Colv = NA)

## case
res1.case <- corX.case - dim1.est
heatmap(res1.case[cor.heat.res1$rowInd, cor.heat.res1$colInd], col = col.heat, Rowv = NA, Colv = NA)

```

### Dimension 3

```{r, include = FALSE}

bar.grad1 <- PrettyBarPlot2(q.grad1[order(dx.grad1),3],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad1[,3] == TRUE, col4grad1, "grey90"),
                            font.size = 0, main = "Scores - gradient 1")

bar.grad2 <- PrettyBarPlot2(q.grad2[order(dx.grad2),3],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad1[,3] == TRUE, col4grad1, "grey90"), 
                            font.size = 0, main = "Scores - gradient 2")

bar.grad3 <- PrettyBarPlot2(q.grad3[order(dx.grad3),3],
                            threshold = 0, 
                            color4bar = ifelse(q.sig.grad1[,3] == TRUE, col4grad1, "grey90"), 
                            font.size = 0, main = "Scores - gradient 3")

```


```{r, fig.show = 'hold', out.width = '50%'}
lxly.out[[3]]

gridExtra::grid.arrange(bar.grad1, bar.grad2, bar.grad3, ncol = 1)

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,3],
               threshold = 0, color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,3] == TRUE, behav.dx$type.col, "grey90"),
               horizontal = FALSE, main = "Scores - behavioural")

dim2.est <- (as.matrix(pls.res$TExPosition.Data$pdq$p[,1:2]) %*% pls.res$TExPosition.Data$pdq$Dd[1:2,1:2] %*% t(as.matrix(pls.res$TExPosition.Data$pdq$q[,1:2])))


cor.heat.res2 <- heatmap(pls.res$TExPosition.Data$X - dim2.est, col = col.heat)

## control
res2.ctrl <- corX.ctrl - dim1.est - dim2.est
heatmap(res2.ctrl[cor.heat.res2$rowInd, cor.heat.res2$colInd], col = col.heat, Rowv = NA, Colv = NA)

## case
res2.case <- corX.case - dim1.est - dim2.est
heatmap(res2.case[cor.heat.res2$rowInd, cor.heat.res2$colInd], col = col.heat, Rowv = NA, Colv = NA)


```

## Back into the brain {.tabset} 

```{r, include = FALSE}
## reorganize data frame
full.pq <- data.frame(grad.dx[,1:3], pls.res$TExPosition.Data$fj[,1:3], colMeans(grad.dat))
colnames(full.pq)[4:7] <- c(paste0("q", c(1:3)), "raw")

full.pq.sig <- full.pq[,4:6]
full.pq.sig[pls.boot$bootRatiosSignificant.j[,1:3] == FALSE] <- 0
full.pq[,8:10] <- full.pq.sig
colnames(full.pq)[8:10] <- c(paste0("qsig", c(1:3)))

pq_for_plot <- full.pq %>% 
  mutate(roi = str_remove(ROI, "_ROI"),
         label = case_when(str_starts(ROI, "L") ~ str_c("lh_", roi),
                           str_starts(ROI, "R_") ~ str_c("rh_", roi)))

gradient_raw_brain <- pq_for_plot %>%
  as.data.frame() %>%
  group_by(gradient) %>%
  ggplot() +
  geom_brain(mapping = aes(fill = raw),
             atlas = glasser) +
  facet_wrap(~gradient, ncol = 1) +
  scale_fill_distiller(palette = "BrBG", limits = c(-1,1), values = c(0, 0.4, 0.5, 0.6, 1)) +
  ggtitle("Gradients")
```

### Dimension 1

```{r, echo = FALSE, warning= FALSE, message='hide', fig.show = 'hold', out.width = '50%'}
gradient_raw_brain

pq_for_plot %>%
  as.data.frame() %>%
  group_by(gradient) %>%
  ggplot() +
  geom_brain(mapping = aes(fill = qsig1),
             atlas = glasser) +
  facet_wrap(~gradient, ncol = 1) +
  scale_fill_distiller(palette = "RdBu", limits = c(-0.7,0.7), values = c(0, 0.45, 0.5, 0.55, 1)) +
  ggtitle("PLSC: Dimension 1") + 
  theme(axis.text.y.left = element_blank(), 
        axis.text.x.bottom = element_blank()) + 
  theme_brain(text.family = "Calibri")

ggplot()

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,1],
               threshold = 0, 
               color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,1] == TRUE, behav.dx$type.col, "grey90"),
               main = "Loadings - behavioural")
```

### Dimension 2

```{r, echo = FALSE, warning= FALSE, message='hide', fig.show = 'hold', out.width = '50%'}
gradient_raw_brain

pq_for_plot %>%
  as.data.frame() %>%
  group_by(gradient) %>%
  ggplot() +
  geom_brain(mapping = aes(fill = qsig2),
             atlas = glasser) +
  facet_wrap(~gradient, ncol = 1) +
  scale_fill_distiller(palette = "RdBu", limits = c(-0.65,0.65), values = c(0, 0.45, 0.5, 0.55, 1)) +
  ggtitle("PLSC: Dimension 2")+ 
  theme(axis.text.y.left = element_blank(), 
        axis.text.x.bottom = element_blank()) + 
  theme_brain(text.family = "Calibri")

ggplot()

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,2],
               threshold = 0, 
               color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,2] == TRUE, behav.dx$type.col, "grey90"),
               main = "Loadings - behavioural")
```

### Dimension 3

```{r, echo = FALSE, warning= FALSE, message='hide', fig.show = 'hold', out.width = '50%'}
gradient_raw_brain

pq_for_plot %>%
  as.data.frame() %>%
  group_by(gradient) %>%
  ggplot() +
  geom_brain(mapping = aes(fill = qsig3),
             atlas = glasser) +
  facet_wrap(~gradient, ncol = 1) +
  scale_fill_distiller(palette = "RdBu", limits = c(-0.7,0.7), values = c(0, 0.45, 0.5, 0.55, 1)) +
  ggtitle("PLSC: Dimension 3")+ 
  theme(axis.text.y.left = element_blank(), 
        axis.text.x.bottom = element_blank()) + 
  theme_brain(text.family = "Calibri")

ggplot()

PrettyBarPlot2(pls.res$TExPosition.Data$fi[,3],
               threshold = 0, 
               color4bar = ifelse(pls.boot$bootRatiosSignificant.i[,3] == TRUE, behav.dx$type.col, "grey90"),
               main = "Loadings - behavioural")
```

## Fancy figures {.tabset}

```{r, include = FALSE}
## organize data - ROI average
spins_mean <- getMeans(spins_grads_num_full[,1:3], spins_grads_num_full$ROI)
grad2plot <- spins_mean
colnames(grad2plot) <- c(paste0("Gradient ", c(1:3)))
grad4grp <- unique(spins_grads_num_full[,c(4,6)])
rownames(grad4grp) <- grad4grp$ROI

## organize data - network mean
net_mean <- getMeans(spins_mean, grad4grp[rownames(spins_mean), "Network"])
colnames(net_mean) <- c(paste0("Gradient ", c(1:3)))

## organize data - ROI & group average
spins_grad_everything <- merge(spins_grads_num_full, sub.dx, by.x = "Subject", by.y = 0)
spins_mean_bygrp<- spins_grad_everything %>% 
  group_by(ROI, Network,diagnostic_group) %>% 
  summarize(across(starts_with("grad"), ~ mean(.x, na.rm = TRUE))) %>%
  as.data.frame %>%
  reshape(idvar = c("ROI", "Network"), timevar = "diagnostic_group", direction = "wide")
grad2plot2 <- spins_mean_bygrp[,c(3:8)]
grad4grp2 <- unique(spins_mean_bygrp[,c(1:2)]) %>% as.data.frame
rownames(grad4grp2) <- grad4grp2$ROI

## organize data - network mean by groups
net_mean_bygrp <- spins_mean_bygrp %>%
  group_by(Network) %>%
  summarize(across(starts_with("grad"), ~ mean(.x, na.rm = TRUE))) %>%
  as.data.frame
rownames(net_mean_bygrp) <- net_mean_bygrp$Network

## get colors
roi.col <- list()
roi.col$oc <- recode(grad4grp$Network, !!!net.col.idx) %>% as.matrix
rownames(roi.col$oc) <- grad4grp$ROI
roi.col$gc <- as.matrix(net.col.idx)

## data for arrows with q
grad_fjxgrad <- data.frame(pls.res$TExPosition.Data$pdq$q[,1:3]*2.5, grad.dx[,c("ROI", "gradient")])
colnames(grad_fjxgrad)[1:3] <- paste0("Dim", c(1:3))
grad_fj_wide <- reshape(grad_fjxgrad, idvar = "ROI", timevar = "gradient", direction = "wide") %>% as.data.frame
grad_fj_wide$Network <- grad4grp[grad_fj_wide$ROI, "Network"]
rownames(grad_fj_wide) <- grad_fj_wide$ROI

net_mean.fj <- getMeans(grad_fj_wide[,2:10], grad_fj_wide$Network)

## color of arrows
arrow.col <- darken(roi.col$gc[rownames(net_mean),], 0.4)
names(arrow.col) <- rownames(net_mean)
```

3D plot of the gradients

```{r, echo = FALSE}
data4plotly <- data.frame(spins_mean, grad4grp[rownames(spins_mean),])
data4plotly.case <- data.frame(spins_mean_bygrp[,3:5], grad4grp2[spins_mean_bygrp$ROI,])
data4plotly.control <- data.frame(spins_mean_bygrp[,6:8], grad4grp2[spins_mean_bygrp$ROI,])
color4plotly <- net.col.idx

## overall average
fig <- plot_ly(data4plotly, x = ~grad1, y = ~grad2, z = ~grad3, color = ~Network, colors = color4plotly, alpha = 0.5, size = 5)  
fig <- fig %>% add_markers()
fig <- fig %>% layout(title = "Overall: mean gradients by ROIs",
                      scene = list(xaxis = list(title = "Gradient 1"),
                                   yaxis = list(title = "Gradient 2"),
                                   zaxis = list(title = "Gradient 3")))
## case
fig.case <- plot_ly(data4plotly.case, x = ~grad1.case, y = ~grad2.case, z = ~grad3.case, color = ~Network, colors = color4plotly, alpha = 0.5, size = 5)  
fig.case <- fig.case %>% add_markers()
fig.case <- fig.case %>% layout(title = "Case: mean gradients by ROIs",
                                scene = list(xaxis = list(title = "Gradient 1"),
                                             yaxis = list(title = "Gradient 2"),
                                             zaxis = list(title = "Gradient 3")))

## control
fig.control <- plot_ly(data4plotly.control, x = ~grad1.control, y = ~grad2.control, z = ~grad3.control, color = ~Network, colors = color4plotly, alpha = 0.5, size = 5)  
fig.control <- fig.control %>% add_markers()
fig.control <- fig.control %>% layout(title = "Control: mean gradients by ROIs",
                                      scene = list(xaxis = list(title = "Gradient 1"),
                                                   yaxis = list(title = "Gradient 2"),
                                                   zaxis = list(title = "Gradient 3")))

## print 3
fig
fig.case
fig.control

```


```{r, echo = FALSE}
### Gradients 1 & 2
G1G2 <- createFactorMap(grad2plot,
                        axis1 = 1, axis2 = 2,
                        col.points = roi.col$oc[rownames(grad2plot),],
                        alpha.points = 0.1, 
                        col.background = "grey96",
                        col.axes = "grey30")

G1G2.mean <- createFactorMap(net_mean,
                             axis1 = 1, axis2 = 2,
                             col.points = roi.col$gc[rownames(net_mean),],
                             col.labels = roi.col$gc[rownames(net_mean),],
                             cex = 4,
                             text.cex = 3,
                             alpha.points = 0.7)

G1G2.mean.control <- createFactorMap(net_mean_bygrp[,5:7],
                                     axis1 = 1, axis2 = 2,
                                     col.points = roi.col$gc[net_mean_bygrp$Network,],
                                     col.labels = roi.col$gc[net_mean_bygrp$Network,],
                                     cex = 2,
                                     text.cex = 3,
                                     alpha.points = 0.7,
                                     col.background = "grey96",
                                     col.axes = "grey30")

G1G2.arrows.dim1 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 2`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim1.grad1, yend = net_mean$`Gradient 2` + net_mean.fj$Dim1.grad2,
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G1G2.arrows.dim2 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 2`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim2.grad1, yend = net_mean$`Gradient 2` + net_mean.fj$Dim2.grad2,
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G1G2.arrows.dim3 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 2`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim3.grad1, yend = net_mean$`Gradient 2` + net_mean.fj$Dim3.grad2,
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)

G1G2.arrows <- annotate("segment", x = net_mean_bygrp$grad1.control, y = net_mean_bygrp$grad2.control, xend = net_mean_bygrp$grad1.case, yend = net_mean_bygrp$grad2.case, 
                        color = "black", alpha = 1, arrow = arrow(length = unit(0.2, "cm"), type = "closed", angle = 30), size = 0.5)

```

```{r, echo = FALSE}
### Gradients 2 & 3
G2G3 <- createFactorMap(grad2plot,
                        axis1 = 2, axis2 = 3,
                        col.points = roi.col$oc[rownames(grad2plot),],
                        alpha.points = 0.1,
                        col.background = "grey96",
                        col.axes = "grey30")

G2G3.mean <- createFactorMap(net_mean,
                             axis1 = 2, axis2 = 3,
                             col.points = roi.col$gc[rownames(net_mean),],
                             col.labels = roi.col$gc[rownames(net_mean),],
                             cex = 4,
                             text.cex = 3,
                             alpha.points = 0.7)

G2G3.mean.control <- createFactorMap(net_mean_bygrp[,5:7],
                                     axis1 = 2, axis2 = 3,
                                     col.points = roi.col$gc[net_mean_bygrp$Network,],
                                     col.labels = roi.col$gc[net_mean_bygrp$Network,],
                                     cex = 2,
                                     text.cex = 3,
                                     alpha.points = 0.7,
                                     col.background = "grey96",
                                     col.axes = "grey30")

G2G3.arrows.dim1 <- annotate("segment", x = net_mean$`Gradient 2`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 2` + net_mean.fj$Dim1.grad2, yend = net_mean$`Gradient 3` + net_mean.fj$Dim1.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G2G3.arrows.dim2 <- annotate("segment", x = net_mean$`Gradient 2`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 2` + net_mean.fj$Dim2.grad2, yend = net_mean$`Gradient 3` + net_mean.fj$Dim2.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G2G3.arrows.dim3 <- annotate("segment", x = net_mean$`Gradient 2`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 2` + net_mean.fj$Dim3.grad2, yend = net_mean$`Gradient 3` + net_mean.fj$Dim3.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)

G2G3.arrows <- annotate("segment", x = net_mean_bygrp$grad2.control, y = net_mean_bygrp$grad3.control, xend = net_mean_bygrp$grad2.case, yend = net_mean_bygrp$grad3.case, 
                        color = "black", alpha = 1, arrow = arrow(length = unit(0.2, "cm"), type = "closed", angle = 30), size = 0.5)
```

```{r, echo = FALSE}
### Gradients 1 & 3
G1G3 <- createFactorMap(grad2plot,
                        axis1 = 1, axis2 = 3,
                        col.points = roi.col$oc[rownames(grad2plot),],
                        alpha.points = 0.1, 
                        col.background = "grey96",
                        col.axes = "grey30")

G1G3.mean <- createFactorMap(net_mean,
                             axis1 = 1, axis2 = 3,
                             col.points = roi.col$gc[rownames(net_mean),],
                             col.labels = roi.col$gc[rownames(net_mean),],
                             cex = 4,
                             text.cex = 3,
                             alpha.points = 0.7)

G1G3.mean.control <- createFactorMap(net_mean_bygrp[,5:7],
                                     axis1 = 1, axis2 = 3,
                                     col.points = roi.col$gc[net_mean_bygrp$Network,],
                                     col.labels = roi.col$gc[net_mean_bygrp$Network,],
                                     cex = 2,
                                     text.cex = 3,
                                     alpha.points = 0.7,
                                     col.background = "grey96",
                                     col.axes = "grey30")


G1G3.arrows.dim1 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim1.grad1, yend = net_mean$`Gradient 3` + net_mean.fj$Dim1.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G1G3.arrows.dim2 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim2.grad1, yend = net_mean$`Gradient 3` + net_mean.fj$Dim2.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G1G3.arrows.dim3 <- annotate("segment", x = net_mean$`Gradient 1`, y = net_mean$`Gradient 3`, xend = net_mean$`Gradient 1` + net_mean.fj$Dim3.grad1, yend = net_mean$`Gradient 3` + net_mean.fj$Dim3.grad3, 
                             color = arrow.col, alpha = 0.8, arrow = arrow(length = unit(0.25, "cm"), type = "closed", angle = 20), size = 1)
G1G3.arrows <- annotate("segment", x = net_mean_bygrp$grad1.control, y = net_mean_bygrp$grad3.control, xend = net_mean_bygrp$grad1.case, yend = net_mean_bygrp$grad3.case, 
                        color = "black", alpha = 1, arrow = arrow(length = unit(0.2, "cm"), type = "closed", angle = 30), size = 0.5)
```

### Group difference

```{r, echo = FALSE, fig.show='hold', out.width='50%'}
G1G2.mean.control$zeMap + G1G2.arrows + ggtitle("From control to case")
G2G3.mean.control$zeMap + G2G3.arrows + ggtitle("From control to case")
G1G3.mean.control$zeMap + G1G3.arrows + ggtitle("From control to case")
```


### Dimension 1

We need to interpret the arrows with cautious, because only the direction and the magnitude are meaningful but not the end point.

```{r, echo = FALSE, fig.show='hold', out.width='50%'}
G1G2$zeMap_background + G1G2$zeMap_dots + G1G2.mean$zeMap_text + G1G2.mean$zeMap_dots + G1G2.arrows.dim1
G2G3$zeMap_background + G2G3$zeMap_dots + G2G3.mean$zeMap_text + G2G3.mean$zeMap_dots + G2G3.arrows.dim1
G1G3$zeMap_background + G1G3$zeMap_dots + G1G3.mean$zeMap_text + G1G3.mean$zeMap_dots + G1G3.arrows.dim1
```

### Dimension 2

We need to interpret the arrows with cautious, because only the direction and the magnitude are meaningful but not the end point.

```{r, echo = FALSE, fig.show='hold', out.width='50%'}
G1G2$zeMap_background + G1G2$zeMap_dots + G1G2.mean$zeMap_text + G1G2.mean$zeMap_dots + G1G2.arrows.dim2
G2G3$zeMap_background + G2G3$zeMap_dots + G2G3.mean$zeMap_text + G2G3.mean$zeMap_dots + G2G3.arrows.dim2
G1G3$zeMap_background + G1G3$zeMap_dots + G1G3.mean$zeMap_text + G1G3.mean$zeMap_dots + G1G3.arrows.dim2
```

### Dimension 3

We need to interpret the arrows with cautious, because only the direction and the magnitude are meaningful but not the end point.

```{r, echo = FALSE, fig.show='hold', out.width='50%'}
G1G2$zeMap_background + G1G2$zeMap_dots + G1G2.mean$zeMap_text + G1G2.mean$zeMap_dots + G1G2.arrows.dim3
G2G3$zeMap_background + G2G3$zeMap_dots + G2G3.mean$zeMap_text + G2G3.mean$zeMap_dots + G2G3.arrows.dim3
G1G3$zeMap_background + G1G3$zeMap_dots + G1G3.mean$zeMap_text + G1G3.mean$zeMap_dots + G1G3.arrows.dim3
```