---
title: "Wrangle SPINS subject info from Lindsay"
output: html_notebook
---

Grabbing SPINS gradients

```{r}
library(tidyverse)
```

## Read in the SPINS big table


```{r}
# read the full grads file
spins_grads <- read_csv("../data/spins_concatenated_full.csv") %>%
  mutate(site = str_sub(`Subject ID`, 5, 7)) %>%
  rename(task = `EA or RS`, 
         subject = `Subject ID`) 

# get lists of the distinct number of subjects
grads_RS_subjects <- spins_grads %>% filter(task=="RS") %>% select(subject) %>% distinct()
grads_EA_subjects <- spins_grads %>% filter(task=="EA") %>% select(subject) %>% distinct()

# remove the full grads file for space
rm(spins_grads)
```



## wrangle clinical data

- figure out exclusions
-we and fd as covariate
-within task

```{r}
RS_QC_exclusion_list <- read_csv("../data/RS_QC_exclusion_list.csv", 
    col_names = "record_id")

EA_QC_exclusion_list <- read_csv("../data/QC_EA_task_exclusion_list_04-26-2021.csv",
    col_names = "record_id")


## contains all SPINS n=557 
spins_behav_data_full_11_20_2020 <- read_csv("../data/spins_behav_data_full_11-20-2020.csv")


# contains n=356 observations
spins_behav_ea_conn_04_29_2021 <- read_csv("../data/spins_behav_ea_conn_04-29-2021.csv")

spins_demo <- spins_behav_data_full_11_20_2020 %>%
  select(record_id, site, scanner, diagnostic_group, demo_sex, demo_age_study_entry, term_early_withdraw) 
  
```


```{r eval = FALSE}
SPINS_fd_by_run <- read_csv("../data/SPINS_fd_by_run_12-14-2020.csv")

fd_score <- SPINS_fd_by_run %>%
    select(-fd_mean.imi_bold, 
           -fd_mean.obs_bold, 
           -`fd_mean.rest_run-1_bold`, 
           -`fd_mean.rest_run-2_bold`, 
           -fd_mean.emp_bold) %>%
  rowwise() %>%
  mutate(fd_max_emp = max(c_across(starts_with("fd_mean.emp")), na.rm = TRUE),
         fd_mean_emp = mean(c_across(starts_with("fd_mean.emp")), na.rm = TRUE)) %>%
  mutate(subject = str_c("sub-", str_sub(record_id, 7,9), str_sub(record_id, 11,15))) 
```

note MRP0147 and MRP0149 have early withdrawal codes - so don't need to worry about these values

```{r}
group_bold <- read_delim("../data/group_bold.tsv", 
    delim = "\t", escape_double = FALSE, 
    trim_ws = TRUE) %>%
  separate(bids_name, 
           into = c("subject", "session", "task", "run"), 
           extra = "merge",
           sep = "_",
           remove = FALSE)

# note - I checked that: percent_fd would not exclude addional people and size_t was acceptable for all scans
fd_info <- group_bold %>%
  filter(task %in% c("task-rest", "task-emp")) %>%
  group_by(subject, session, task) %>%
  summarise(n_scans = n(),
            fd_mean = mean(fd_mean)) %>%
  pivot_wider(id_cols = c("subject", "session"),
              names_from = task,
              values_from = c(n_scans, fd_mean))
```












Note: teh weird "fd_mean.emp" and rest runs one and two come from pilots or early withdrawal participants - so we can drop those columns

We are also not using the imiate or observe tasks for this

```{r}
spins_demo %>%
  mutate(subject = str_c("sub-", str_sub(record_id,7,9), str_sub(record_id,11)))
```



```{r}
spins_demo_withAC <- spins_demo %>%
  mutate(subject = str_c("sub-", str_sub(record_id,7,9), str_sub(record_id,11))) %>%
  left_join(fd_info, by = "subject") %>%
  mutate(has_RS_grads = subject %in% grads_EA_subjects$subject,
         has_EA_grads = subject %in% grads_RS_subjects$subject,
         RS_QC_exclude = record_id %in% RS_QC_exclusion_list$record_id,
         EA_QC_exclude = record_id %in% EA_QC_exclusion_list$record_id,
         RS_mo_exclude = `fd_mean_task-rest` > 0.5,
         EMP_mo_exclude = `fd_mean_task-emp` > 0.5,
         withdrawn_exclude = !(is.na(term_early_withdraw))) %>%
  rowwise() %>%
  mutate(exclude_MRI = any(c_across(ends_with("exclude")), na.rm = TRUE))
```

| code | interpretation |
|---|---|
| 1 | didn't do MRI | 
| 2 | late medical exclusion (medication change or high blood sugar) |
| 3 | unknown (odd bc some have MR data) |
| 6 | unknown (but none have scan data) |
| 8 | withdrawn by PI - late medical exclusion |
| 9 | unknown (but no EA data) |
| 10 | did not complete full protocol |

In general - we might as well remove anyone with a withdrawn code from the study (even if they have MR data)

```{r}
spins_demo_withAC %>%
  mutate(across(.cols = c("site", "scanner", "diagnostic_group", "demo_sex"), 
                .fns = factor)) %>%
  summary()
```



```{r}
spins_demo_withAC %>%
  count(exclude_MRI, has_EA_grads, has_RS_grads)
```
```{r}
spins_demo_withAC %>%
  filter(exclude_MRI == FALSE, has_EA_grads == FALSE, has_RS_grads == FALSE)
```

```{r}
spins_behav_ea_conn_04_29_2021 %>% names()
```







```{r}
spins_demo_withAC %>%
  filter(exclude_MRI==FALSE, 
         has_EA_grads==TRUE, 
         has_RS_grads==TRUE) %>%
  write_csv('../data/spins_subject_info_for_grads_2021-07-29.csv')
```

```{r}
demo_vars = c('record_id', 'site', 'scanner', 'diagnostic_group', 'demo_sex', 'demo_age_study_entry', 'term_early_withdraw')
clinical_vars = c("scog_tasit1_total", "scog_tasit2_total", "scog_tasit3_total", 
                  "np_composite_tscore",  
                  "np_domain_tscore_process_speed",   "np_domain_tscore_att_vigilance",
                  "np_domain_tscore_work_mem","np_domain_tscore_verbal_learning",
                  "np_domain_tscore_visual_learning",
                  "np_domain_tscore_reasoning_ps","np_domain_tscore_social_cog",
                  "bprs_factor_pos_symp", "bprs_factor_total", "bprs_factor_neg_symp",
                  "qls_total", "sans_total_sc", "bsfs_total")

lol_spins_behav <- spins_behav_ea_conn_04_29_2021 %>%
  select(all_of(demo_vars), all_of(clinical_vars)) %>%
  mutate(subject = str_c("sub-", str_sub(record_id,7,9), str_sub(record_id,11))) %>%
  left_join(fd_info, by = "subject") %>%
  mutate(has_RS_grads = subject %in% grads_EA_subjects$subject,
         has_EA_grads = subject %in% grads_RS_subjects$subject,
         RS_QC_exclude = record_id %in% RS_QC_exclusion_list$record_id,
         EA_QC_exclude = record_id %in% EA_QC_exclusion_list$record_id,
         RS_mo_exclude = `fd_mean_task-rest` > 0.5,
         EA_mo_exclude = `fd_mean_task-emp` > 0.5,
         withdrawn_exclude = !(is.na(term_early_withdraw))) %>%
  rowwise() %>%
  mutate(exclude_MRI = any(c_across(ends_with("exclude")), na.rm = TRUE))
```

```{r}
lol_spins_behav %>%
  filter(exclude_MRI==TRUE)
```
```{r}
lol_spins_behav %>% 
    mutate(across(.cols = c("site", "scanner", "diagnostic_group", "demo_sex"), 
                .fns = factor)) %>%
  filter(exclude_MRI==FALSE) %>%
  summary()
```


One weird observation - CMP0218 - CMP0220 missing their MRI QC bold scores 


```{r}
lol_spins_behav %>%
  filter(exclude_MRI==FALSE) %>%
  write_csv('../data/spins_lolivers_subject_info_for_grads_2022-02-02.csv')
```


Notes from lindsay  
  
So the current df is here: /projects/loliver/SPINS_PLS_Conn/data/processed/spins_behav_ea_conn_04-29-2021.csv
6:27
However, I also excluded ZHP_0110 and ZHP_0172 as they did not complete the soc cog tasks
6:27
So that df includes 356, but my current N is 354

The following participants were missing task-rest cleaned files:
sub-CMP0180
sub-CMP0182
sub-CMP0191
sub-CMP0196
sub-CMP0198
sub-CMP0207
sub-CMP0213
sub-ZHH0034

The following participants were missing task-emp background files:
sub-CMH0025
sub-CMH0044
sub-CMH0095
sub-CMH0138
sub-MRC0075
sub-MRP0077
sub-MRP0147
sub-MRP0149
sub-MRP0151
sub-MRP0161
sub-ZHH0038
sub-ZHP0061
sub-ZHP0086




