---
title: "Wrangle SPINS subject info from Lindsay"
output: html_notebook
---

Grabbing SPINS gradients

```{r}
library(tidyverse)
```

## Read in the SPINS big table


```{r}

spins_grads <- read_csv("../data/spins_concatenated_full.csv")
spins_grads <- spins_grads %>%
  mutate(site = str_sub(`Subject ID`, 5, 7)) %>%
  rename(task = `EA or RS`, 
         subject = `Subject ID`)
```


## wrangle clinical data

- figure out exclusions
-we and fd as covariate
-within task

```{r}
RS_QC_exclusion_list <- read_csv("../data/RS_QC_exclusion_list.csv", 
    col_names = "record_id")

EA_QC_exclusion_list <- read_csv("../data/QC_EA_task_exclusion_list_04-26-2021.csv",
    col_names = "record_id")

spins_behav_data_full_11_20_2020 <- read_csv("../data/spins_behav_data_full_11-20-2020.csv")

spins_demo <- spins_behav_data_full_11_20_2020 %>%
  select(record_id, site, scanner, diagnostic_group, demo_sex, demo_age_study_entry, term_early_withdraw) 
  
```
```{r}
SPINS_fd_by_run <- read_csv("../data/SPINS_fd_by_run_12-14-2020.csv")
```

note MRP0147 and MRP0149 have early withdrawal codes - so don't need to worry about these values

```{r}
SPINS_fd_by_run %>%
  drop_na(fd_mean.task)
```

```{r}
RS_QC_exclusion_list %>%
  mutate(subject = str_c("sub-", str_sub(subid, 7,9), str_sub(subid, 11,15))) 
```

```{r}
grads_RS_subjects <- spins_grads %>% filter(task=="RS") %>% select(subject) %>% distinct()
grads_EA_subjects <- spins_grads %>% filter(task=="EA") %>% select(subject) %>% distinct()
```


```{r} 
SPINS_fd_by_run %>%
  mutate(subject = str_c("sub-", str_sub(record_id, 7,9), str_sub(record_id, 11,15))) %>%
  filter(`fd_mean.emp_run-1_bold` > 0.5)
```

Note: teh weird "fd_mean.emp" and rest runs one and two come from pilots or early withdrawal participants - so we can drop those columns

We are also not using the imiate or observe tasks for this



```{r} 
fd_score <- SPINS_fd_by_run %>%
    select(-fd_mean.imi_bold, 
           -fd_mean.obs_bold, 
           -`fd_mean.rest_run-1_bold`, 
           -`fd_mean.rest_run-2_bold`, 
           -fd_mean.emp_bold) %>%
  rowwise() %>%
  mutate(fd_max_emp = max(c_across(starts_with("fd_mean.emp")), na.rm = TRUE),
         fd_mean_emp = mean(c_across(starts_with("fd_mean.emp")), na.rm = TRUE)) %>%
  mutate(subject = str_c("sub-", str_sub(record_id, 7,9), str_sub(record_id, 11,15))) 

```
```{r}
spins_demo_withAC <- spins_demo %>%
  left_join(fd_score, by = "record_id") %>%
  mutate(has_RS_grads = subject %in% grads_EA_subjects$subject,
         has_EA_grads = subject %in% grads_RS_subjects$subject,
         RS_QC_exclude = record_id %in% RS_QC_exclusion_list$record_id,
         EA_QC_exclude = record_id %in% EA_QC_exclusion_list$record_id,
         RS_mo_exclude = `fd_mean.rest_bold` > 0.5,
         RS_mo_exclude = `fd_mean_emp` > 0.5,
         withdrawn_exclude = !(is.na(term_early_withdraw))) %>%
  rowwise() %>%
  mutate(exclude_MRI = any(c_across(ends_with("exclude")), na.rm = TRUE))
```

| code | interpretation |
|---|---|
| 1 | didn't do MRI | 
| 2 | late medical exclusion (medication change or high blood sugar) |
| 3 | unknown (odd bc some have MR data) |
| 6 | unknown (but none have scan data) |
| 8 | withdrawn by PI - late medical exclusion |
| 9 | unknown (but no EA data) |
| 10 | did not complete full protocol |

In general - we might as well remove anyone with a withdrawn code from the study (even if they have MR data)

```{r}
spins_demo_withAC %>%
  summary
```


```{r}
spins_demo_withAC %>%
  count(exclude_MRI, has_EA_grads, has_RS_grads)
```

```{r}
write_csv(../data/spins)
```


```{r}
spins_behav_ea_conn_04_29_2021 <- read_csv("../data/spins_behav_ea_conn_04-29-2021.csv")

```


```{r}
spins_demo_withAC %>%
  anti_join(spins_behav_ea_conn_04_29_2021, by = "record_id") %>%
  filter(exclude_MRI==FALSE, 
         has_EA_grads==TRUE, has_RS_grads==TRUE) 
  
```





```{r}
spins_demo_withAC %>%
  filter(exclude_MRI==FALSE, 
         has_EA_grads==TRUE, 
         has_RS_grads==TRUE) %>%
  write_csv('../data/spins_subject_info_for_grads_2021-07-29.csv')
```

```{r}
lol_spins_behav <- spins_behav_ea_conn_04_29_2021 %>%
  select(record_id, site, scanner, diagnostic_group, demo_sex, demo_age_study_entry, term_early_withdraw) %>%
  left_join(fd_score, by = "record_id") %>%
  mutate(has_RS_grads = subject %in% grads_EA_subjects$subject,
         has_EA_grads = subject %in% grads_RS_subjects$subject,
         RS_QC_exclude = record_id %in% RS_QC_exclusion_list$record_id,
         EA_QC_exclude = record_id %in% EA_QC_exclusion_list$record_id,
         RS_mo_exclude = `fd_mean.rest_bold` > 0.5,
         RS_mo_exclude = `fd_mean_emp` > 0.5,
         withdrawn_exclude = !(is.na(term_early_withdraw))) %>%
  rowwise() %>%
  mutate(exclude_MRI = any(c_across(ends_with("exclude")), na.rm = TRUE))
```

```{r}
lol_spins_behav %>%
  filter(exclude_MRI==TRUE)
```
```{r}
lol_spins_behav %>% summary()
```
```{r}
lol_spins_behav %>%
  write_csv('../data/spins_lolivers_subject_info_for_grads_2021-07-29.csv')
```


Notes from lindsay  
  
So the current df is here: /projects/loliver/SPINS_PLS_Conn/data/processed/spins_behav_ea_conn_04-29-2021.csv
6:27
However, I also excluded ZHP_0110 and ZHP_0172 as they did not complete the soc cog tasks
6:27
So that df includes 356, but my current N is 354

The following participants were missing task-rest cleaned files:
sub-CMP0180
sub-CMP0182
sub-CMP0191
sub-CMP0196
sub-CMP0198
sub-CMP0207
sub-CMP0213
sub-ZHH0034

The following participants were missing task-emp background files:
sub-CMH0025
sub-CMH0044
sub-CMH0095
sub-CMH0138
sub-MRC0075
sub-MRP0077
sub-MRP0147
sub-MRP0149
sub-MRP0151
sub-MRP0161
sub-ZHH0038
sub-ZHP0061
sub-ZHP0086




