---
title: "Wrangle SPINS gradients with GSR"
output:
  html_document:
    df_print: paged
---

Grabbing SPINS gradients

```{r}
library(tidyverse)
library(ggseg)
library(ggsegGlasser)
library(broom)
```

## Read in the SPINS big table


```{r}

spins_grads_wGSR <- read_csv("../gsr_spins_concat_full")
# spins_grads <- spins_grads %>%
#   mutate(site = str_sub(`Subject`, 5, 7)) %>%
#   rename(task = `EA or RS`, 
#          subject = `Subject ID`)
spins_grads_noGSR <- read_csv("../spins_concat_full")

spins_grads <- bind_rows("GSR" = spins_grads_wGSR,
                     "noGSR" = spins_grads_noGSR,
                     .id = "GSR")

rm(spins_grads_noGSR, spins_grads_wGSR)
```


## read subject data


```{r}
lol_spins_behav <- 
  read_csv('../data/spins_lolivers_subject_info_for_grads_2022-02-02.csv')

```
```{r}
names(lol_spins_behav)
```

## grab some network colours

```{r}
networks <- read_delim("../networks.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  select(NETWORK, NETWORKKEY, RED, GREEN, BLUE, ALPHA) %>%
  distinct() %>%
  add_row(NETWORK = "Subcortical", NETWORKKEY = 13, RED = 0, GREEN=0, BLUE=0, ALPHA=255) %>%
  mutate(hex = rgb(RED, GREEN, BLUE, maxColorValue = 255)) %>%
  arrange(NETWORKKEY)

networks
```


## fitting models

Models

- model includes DX, sex, age, scanner, fd?

- what about between task - 
  - do paired ttest?
  
  
So the current df is here: /projects/loliver/SPINS_PLS_Conn/data/processed/spins_behav_ea_conn_04-29-2021.csv
6:27
However, I also excluded ZHP_0110 and ZHP_0172 as they did not complete the soc cog tasks
6:27
So that df includes 356, but my current N is 354

The following participants were missing task-rest cleaned files:
sub-CMP0180
sub-CMP0182
sub-CMP0191
sub-CMP0196
sub-CMP0198
sub-CMP0207
sub-CMP0213
sub-ZHH0034

The following participants were missing task-emp background files:
sub-CMH0025
sub-CMH0044
sub-CMH0095
sub-CMH0138
sub-MRC0075
sub-MRP0077
sub-MRP0147
sub-MRP0149
sub-MRP0151
sub-MRP0161
sub-ZHH0038
sub-ZHP0061
sub-ZHP0086

## 

```{r}
overall_grand_mean_grads <- spins_grads %>%
  mutate(Network = factor(Network, levels = networks$NETWORK)) %>%
  filter(Subject %in% lol_spins_behav$subject) %>%
  group_by(Task, ROI, Network, GSR) %>%
  summarise_at(vars(grad1:grad3), mean)

#grand_mean_grads
```

```{r}
## a very convoluted way to pull networks of interest
networks_of_interest <- overall_grand_mean_grads %>%
  group_by(Network) %>%
  summarise_at(vars(grad1:grad3), mean) %>%
  pivot_longer(starts_with("grad"),
               names_to = "grad_name", 
               values_to = "grad_value") %>%
  filter(abs(grad_value) > 0.2) %>%
  distinct(Network) %>%
  pull(Network) %>% as.character()
  
```


```{r}
grand_mean_grads <- spins_grads %>%
  mutate(Network = factor(Network, levels = networks$NETWORK)) %>%
  filter(Subject %in% lol_spins_behav$subject) %>%
  group_by(Site, Task, ROI, Network, GSR) %>%
  summarise_at(vars(grad1:grad3), mean)

#grand_mean_grads
```

```{r}
glasser_atlas <-
  ggsegGlasser::glasser %>% as.tibble() %>% select(hemi, region, side, label)
```


```{r}
overall_grads_for_plot <- overall_grand_mean_grads %>% 
  mutate(roi = str_remove(ROI, "_ROI"),
         label = case_when(str_starts(ROI, "L") ~ str_c("lh_", roi),
                           str_starts(ROI, "R_") ~ str_c("rh_", roi))) %>%
  
  filter(Task == "RS", GSR == "GSR") %>%
  select(starts_with("grad"), label)
```

```{r}
overall_grads_for_plot %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = grad1),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```
```{r}
overall_grads_for_plot %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = grad2),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```

```{r}
overall_grads_for_plot %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = grad3),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```

```{r fig.height=4, fig.width=12}
grand_mean_grads %>%
  ggplot(aes(x = grad1, y = grad2, colour = Network)) +
  geom_point() +
  scale_color_manual(values = networks$hex) +
  facet_grid(Task*GSR ~ Site)
```
```{r fig.height=4, fig.width=12}
grand_mean_grads %>%
  ggplot(aes(x = grad1, y = grad3, colour = Network)) +
  geom_point() +
  scale_color_manual(values = networks$hex) +
  facet_grid(Task*GSR ~ Site)
```
```{r fig.height=4, fig.width=12}
grand_mean_grads %>%
  ggplot(aes(x = grad2, y = grad3, colour = Network)) +
  geom_point() +
  scale_color_manual(values = networks$hex) +
  facet_grid(Task*GSR ~ Site)
```

From age paper:
Each axis of this 3D space was defined by the values along the first three gradients. Within network dispersion was quantified as sum squared Euclidean distance of network nodes to the network centroid at individual level. Between network dispersion was calculated as the Euclidean distance between network centroids.

They also did linear models on nodal values from the first three gradients

## calculating centroid and within network dispersion

```{r}
## calculate the network centers
subject_level_centers <- spins_grads %>%
  filter(Subject %in% lol_spins_behav$subject) %>%
  group_by(Site, Task, Network, Subject, GSR) %>%
  summarise_at(c("grad1", "grad2", "grad3"), mean)
  
subject_within_dispersion <- subject_level_centers %>%
  
  ## combine with orig spins grads
  inner_join(spins_grads, by = c("Site", "Task", "Network", "Subject", "GSR"), suffix = c("_center", "")) %>%
  
  ## calculate squared euclidian distance from the network centroid  
  ungroup() %>%
  mutate(roi_distance_sq = (grad1_center - grad1)**2 + (grad2_center - grad2)**2 + (grad3_center - grad3)**2) %>% 
  
  ## dispersion as the sum of the squared distances within network and participant
  group_by(Site, Task, GSR, Network, Subject) %>%
  summarise(within_dispersion = sum(roi_distance_sq))
  
```

## calculating between network dispersion

Between network dispersion was calculated as the Euclidean distance between network centroids.

```{r}
library(igraph)

## assumes that the network_centroids contains columns "Network", "grad1", "grad2", "grad3"
calc_between_distances <- function(network_centoids) {
  ## grab the unique network node names
  network_names <- unique(network_centoids$Network)

  ## create a graph dataframe of edges
  result <- make_full_graph(length(network_names), directed = FALSE) %>% as_data_frame()
  
  for (i in 1:nrow(result)) {
    to_name <- network_names[result$to[i]]
    from_name <- network_names[result$from[i]]
   
    result$to_name[i] <- to_name
    result$from_name[i] <- from_name

  res_df <- test_centroids %>%
    filter(Network == to_name | Network == from_name) %>%
    mutate_at(vars(c("grad1", "grad2", "grad3")), function(x) {(x - lag(x))**2}) %>%
    mutate(eucl_dist = sqrt(grad1 + grad2 + grad3))

  result$eucl_dist[i] <- res_df$eucl_dist[2]

  }

  return(result %>% select(to_name, from_name, eucl_dist))
}
```


```{r  eval = FALSE}
between_network <- subject_level_centers %>%
  ungroup() %>%
  group_by(Site, Task, Subject, GSR) %>%
  nest() %>%
  mutate(distances = map(data, calc_between_distances)) 

```

```{r eval=FALSE}
between_network <- subject_level_centers %>%
  ungroup() %>%
  group_by(Site, Task, Subject, GSR) %>%
  nest() %>%
  slice(1)

between_network$data
```


## testing for EMP vs REST differences

```{r}
library(broom)
subject_level_centers %>%
  filter(GSR=="GSR") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  ungroup() %>%
  group_by(Network, GSR) %>%
  do(tidy(aov(grad1 ~ Task + fd_mean + scanner + Error(Subject), data = .))) %>%
  filter(term == "Task") %>%
  arrange(GSR) %>%
  knitr::kable()
```
## fit case-control and age differences in gradient dispersion

```{r}
results_subject_within_dispersion <- subject_within_dispersion %>%
  filter(GSR=="GSR", Task=="RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  group_by(Network, Task, GSR) %>%
  do(tidy(lm(within_dispersion ~ diagnostic_group + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

results_subject_within_dispersion %>%
  filter(term %in% c("diagnostic_groupcontrol", 
                     "demo_age_study_entry")) %>%
  arrange(GSR) %>%
  filter(Network %in% networks_of_interest) %>%
  knitr::kable()
```
```{r}
results_subject_within_dispersion %>%
  filter(term == "demo_age_study_entry") %>%
  knitr::kable()
```
```{r fig.height=4, fig.width=8}
library(ggridges)
library(forcats)
subject_level_centers %>%
  filter(GSR=="GSR") %>%
  filter(Network %in% networks_of_interest) %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  ggplot(aes(x = grad1, y = fct_reorder(Network, grad1, .fun = median))) +
  geom_density_ridges() +
  # scale_fill_manual(values = networks$hex) +
  facet_grid(diagnostic_group~Task)
```
```{r fig.height=4, fig.width=8}
subject_level_centers %>%
  filter(GSR=="GSR") %>%
  filter(Network %in% networks_of_interest) %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  ggplot(aes(x = grad2, y = fct_reorder(Network, grad2, .fun = median))) +
  geom_density_ridges() +
  # scale_fill_manual(values = networks$hex) +
  facet_grid(diagnostic_group~Task)
```

```{r fig.height=4, fig.width=8}
subject_level_centers %>%
  filter(GSR=="GSR") %>%
  filter(Network %in% networks_of_interest) %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  ggplot(aes(x = grad3, y = fct_reorder(Network, grad3, .fun = median))) +
  geom_density_ridges() +
  # scale_fill_manual(values = networks$hex) +
  facet_grid(diagnostic_group~Task*GSR)
```
## fit network locations against case-control

```{r}
net_centroids_results <- subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%

  pivot_longer(starts_with("grad"), names_to = "grad_num", values_to = "grad_value") %>%
    ungroup() %>%
  group_by(Network, Task, GSR, grad_num) %>%
  do(tidy(lm(grad_value ~ diagnostic_group + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

net_centroids_results %>%
  filter(term == "diagnostic_groupcontrol") %>%
  filter(Network %in% networks_of_interest, GSR == "GSR") %>%
  knitr::kable()
```



```{r}
all_roi_results <- spins_grads %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  pivot_longer(grad1:grad3, names_to = "grad_num", values_to = "grad_value") %>%
    ungroup() %>%
  group_by(Network, Task, , GSR, grad_num, ROI) %>%
  do(tidy(lm(grad_value ~ diagnostic_group + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))
  
#all_roi_results %>% write_csv('ROIwise_SPINS_SSDvsHC_result_wGSR.csv')
```
```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol",
         grad_num == "grad1",
         Task == "RS") %>%
  mutate(roi = str_remove(ROI, "_ROI"),
         label = case_when(str_starts(ROI, "L") ~ str_c("lh_", roi),
                           str_starts(ROI, "R_") ~ str_c("rh_", roi))) %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = statistic),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```
```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol",
         grad_num == "grad2",
         Task == "RS") %>%
  mutate(roi = str_remove(ROI, "_ROI"),
         label = case_when(str_starts(ROI, "L") ~ str_c("lh_", roi),
                           str_starts(ROI, "R_") ~ str_c("rh_", roi))) %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = statistic),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```

```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol",
         grad_num == "grad3",
         Task == "RS") %>%
  mutate(roi = str_remove(ROI, "_ROI"),
         label = case_when(str_starts(ROI, "L") ~ str_c("lh_", roi),
                           str_starts(ROI, "R_") ~ str_c("rh_", roi))) %>%
  filter(Network != "Subcortical") %>%
  filter(ROI != "L_10pp_ROI") %>%
  as.data.frame() %>%
  ggplot() +
  geom_brain(mapping = aes(fill = statistic),
        atlas = glasser) +
  scale_fill_distiller(type = "div")
```


```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol", p_FDR < 0.05)
```

```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol", p_FDR < 0.05) %>%
  count(Task,GSR, grad_num) %>%
  arrange(Task, GSR, grad_num)
```

```{r}
all_roi_results %>%
  filter(term == "diagnostic_groupcontrol", p_FDR < 0.05) %>%
  count(Task, GSR, grad_num, Network) %>%
  pivot_wider(names_from = grad_num, values_from = n)
  
```

# Mass net centres vs clinical vars

```{r}
symptom_vars <- c("bprs_factor_total", "bprs_factor_neg_symp", 
                  "bprs_factor_neg_symp",
                   "sans_total_sc")

net_centroids_symptoms_results <- subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  pivot_longer(starts_with("grad"), names_to = "grad_num", values_to = "grad_value") %>%
  pivot_longer(all_of(symptom_vars), names_to = "symptom_scale", values_to = "symp_value") %>%
    ungroup() %>%
  group_by(Network, Task, GSR, grad_num, symptom_scale) %>%
  do(tidy(lm(grad_value ~ symp_value + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

net_centroids_symptoms_results %>%
  filter(term == "symp_value") %>%
  filter(Network %in% networks_of_interest, GSR == "GSR") %>%
  arrange(symptom_scale, grad_num) %>%
  knitr::kable()
```

### plot sans total score agains auditory grad2

```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Auditory") %>%
  ggplot(aes(x = diagnostic_group, y = grad2)) +
    geom_boxplot(outlier.shape = NULL) +
    geom_jitter(aes(colour = diagnostic_group)) +
  labs(y = "Auditory Network Gradient 2 value (negative is more defined from visual)")
```

```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Auditory") %>%
  ggplot(aes(x = sans_total_sc, y = grad2)) +
    geom_point(aes(colour = diagnostic_group)) + 
    geom_smooth(method = "lm")
```
```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Auditory") %>%
  ggplot(aes(x = sans_total_sc, y = grad2)) +
    geom_point(aes(colour = diagnostic_group)) + 
    geom_smooth(method = "lm") +
  facet_wrap(~site)
```


---> ROIwise SANS by grad2

## neurocog and functioning my grads

```{r}
functioning_vars <- c("qls_total", "bsfs_total")

net_centroids_functioning_results <- subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  pivot_longer(starts_with("grad"), names_to = "grad_num", values_to = "grad_value") %>%
  pivot_longer(all_of(functioning_vars), names_to = "func_scale", values_to = "func_value") %>%
    ungroup() %>%
  group_by(Network, Task, GSR, grad_num, func_scale) %>%
  do(tidy(lm(grad_value ~ func_value + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

net_centroids_functioning_results %>%
  filter(term == "func_value") %>%
  filter(Network %in% networks_of_interest, GSR == "GSR") %>%
  arrange(func_scale, grad_num) %>%
  knitr::kable()
```

```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Visual1") %>%
  ggplot(aes(x = bsfs_total, y = grad1, colour = diagnostic_group)) +
    geom_point() + 
    geom_smooth(method = "lm")
```

```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Auditory") %>%
  ggplot(aes(x = bsfs_total, y = grad2, colour = diagnostic_group)) +
    geom_point() + 
    geom_smooth(method = "lm")
```
```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Visual2") %>%
  ggplot(aes(x = qls_total, y = grad2, colour = diagnostic_group)) +
    geom_point() + 
    geom_smooth(method = "lm")
```

## Just the two vars for the SOBP abstract

```{r}
functioning_vars <- c("sans_total_sc", "bsfs_total")

net_centroids_functioning_results <- subject_level_centers %>%
  filter(GSR=="GSR", 
         Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  pivot_longer(starts_with("grad"), names_to = "grad_num", values_to = "grad_value") %>%
  pivot_longer(all_of(functioning_vars), names_to = "func_scale", values_to = "func_value") %>%
    ungroup() %>%
  group_by(Network, Task, GSR, grad_num, func_scale) %>%
  do(tidy(lm(grad_value ~ func_value + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

net_centroids_functioning_results %>%
  filter(term == "func_value") %>%
  filter(Network %in% networks_of_interest, GSR == "GSR") %>%
  arrange(func_scale, grad_num) %>%
  knitr::kable()
```

## Gradient values against cognition



```{r}


net_centroids_functioning_results <- subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  mutate( fd_mean = if_else(Task=="EA", `fd_mean_task-emp`, `fd_mean_task-rest`)) %>%
  pivot_longer(starts_with("grad"), names_to = "grad_num", values_to = "grad_value") %>%
  pivot_longer(scog_tasit1_total:np_domain_tscore_social_cog, names_to = "cog_scale", values_to = "cog_value") %>%
    ungroup() %>%
  group_by(Network, Task, GSR, grad_num, cog_scale) %>%
  do(tidy(lm(grad_value ~ cog_value  + demo_age_study_entry + demo_sex + fd_mean + scanner, data = .))) %>%
  ungroup() %>%
  group_by(term) %>%
  mutate(p_FDR = p.adjust(p.value, method = "fdr"))

net_centroids_functioning_results %>%
  filter(term == "cog_value") %>%
  filter(Network %in% networks_of_interest, GSR == "GSR") %>%
  arrange(cog_scale, grad_num) %>%
  knitr::kable()
```
```{r}
subject_level_centers %>%
  filter(GSR=="GSR", Task == "RS") %>%
  inner_join(lol_spins_behav, by = c("Subject" = "subject")) %>%
  filter(Network == "Visual2") %>%
  ggplot(aes(x = np_composite_tscore, y = grad2, colour = diagnostic_group)) +
    geom_point() + 
    geom_smooth(method = "lm")
```
Therefore - the same relationship holds with cognition  - but seems to be mostly driven by the diagnostic group differences

```{r}
lol_spins_behav %>% 
  ggplot(aes(x = sans_total_sc, y = `fd_mean_task-rest`)) +
  geom_point()
```

## manova
## correlational distance

  
## 